---
title: "PUBG Finish Placement Prediction"
output: html_notebook

---

# Introduction

Player unknowns battleground competition on kaggle.

The aim is to predict the winner of the 100-person battle royale shooter game.


# Exploratory Data Analysis

```{r load-packages, include=FALSE, warning=FALSE}

# library(xgboost)
# library(magrittr)
# library(tidyverse)
# library(corrplot)
# library(GGally)
pacman::p_load(tidyverse, xgboost, magrittr, corrplot, GGally)

options(scipen = 999)


source("pubg-model-helper.R")

```

```{r read-data, include=FALSE, warning=FALSE}

train <- read.csv("../data/train_V2.csv")
test <- read.csv("../data/test_V2.csv")

```


First let's have a quick look at the data itself, to check for cleanliness.

```{r summary-statistics, include=FALSE}

# show how many NAs in each column
lapply(train, function(x) sum(is.na(x))) # apparently our data is clean!

# percentage of unique values in each column
lapply(train, function(x) sprintf("number of unique observations by column: %4f%%",
                                  length(unique(x))/length(x)))

# what does our data look like?
head(train)
train %>% names

```

Some thoughts around all the different ID columns:
* How does Id relate to groupId?
* Are shared groupId values the same team?
* How many different matches are there?

```{r distinct-counts, include=FALSE}

count_distinct_col <- function(data, key) {
  data %>%
    select(key) %>%
    distinct %>%
    nrow
}

# can 1 person play multiple games?

# different game modes
# what's fpp???
train %>% count(matchType)

# how many matches?
count_distinct_col(train, "matchId")

# how many groups?
train %>%
  group_by(groupId) %>%
  summarise(count = n())

count_distinct_col(train, "groupId")

# average group size?
train %>%
  group_by(groupId) %>%
  summarise(count = n()) %>%
  summarise(avg_group_size = mean(count))


```

```{r some-cleaning, include=FALSE}

# remove NA winPlacePerc player
train %<>% filter(!is.na(winPlacePerc))

```


Might be worth splitting the data into different groups: one group for purely solo players, one group for duos and squads, and one group for special game modes?

Check if special modes are solo or duo/squad and merge them onto the other 2 sets

Keep in mind that in solo matches you can still have teamKills. These are suicides or accidental deaths from what I can tell

```{r split-data-solo-group}
# TODO: look at how group statistics are treated in solo matches, and vice versa
# TODO: split the data by group
# consider teamKills in solo?
# even in solo mode, some people have a teamKill? what does this mean?
# SUICIDES AND ACCIDENTAL DEATHS!

train %>% 
  filter(matchType == "solo",
         teamKills == 1) %>%
  head

train %>%
  filter(matchType == "crashfpp") %>%
  count(groupId, matchId) %>% head(20)

train %>% 
  filter(groupId == "01d8a1a3ebf59f", matchId == "c392b6578f915b")
```

```{r split-data-by-group-size}

train %>% count(matchType)

# solo-games only

# duo-games only

# squad-games only

# team-games only (combined duo and squad)

```


## Univariate visualisations

> Probably should re-do this after splitting the data into solo and group games

As someone that has played battle royales, I think that a correlation might exist between the number of kills and the win placement. Generally more skilled players in battle royale 'hunt' for kills, while lesser skilled players shy away from gunfights

Other variables that are interesting to look at:
* walkDistance
* weaponsAcquired (might be correlated with number of kills)
* winPlacePerc (target variable)
* damageDealt
* headShotKills (probably correlated with damageDealt. Does mean better aim)


```{r correlation-matrix, echo=FALSE}

# TODO: something about winPlacePerc is weird

corrplot(cor(train %>% select(-c(Id, groupId, matchId, matchType, winPlacePerc))),
         method = "color",
         type = "upper", number.cex = 0.7,
         tl.col = "black", tl.srt = 90, tl.cex = 0.75,
         diag = FALSE
         )

```

Initial findings/things to look into from the correlation plot:

* Something weird about winPlacePerc
* _Very strong_ negative correlation between rankPoints and killPoints, and winPoints and rankPoints. Investigate these relationships
* damageDealt and DBNOs strongly positively correlation
* killPoints and winPoints are strongly correlation
* numGroups and maxPlace strongly positively correlated. Why?
* killPlace negatively correlates strongly with a number of predictors. These are kills, killStreaks, longestKill, walkDistance, and weaponsAcquired

```{r kill-stat-density, echo=FALSE}

# DATA IS WAY TOO BIG TO RUN A GGPPAIRS

kill_pred_incides <- match(c("killPlace", "kills", "killStreaks", "longestKill", "walkDistance", "weaponsAcquired"), names(train))

ggpairs(train, columns = kill_pred_incides, title = "Kill statistics investigation")

```

```{r target-var-explore, include=FALSE}
# what am i actually predicting?
# predicting the probability that the player will win
# a 0 = player will absolutely lose
# a 1 = player will absolutely win
# are there any 0s and 1s in our data?
train %>%
  select(winPlacePerc) %>% head

train %>%
  summary()

# why is there an NA in winPlacePerc?
# TODO: REMOVE
train %>% filter(is.na(winPlacePerc))

# plot winPoints against winPlacePerc
# this is shit - fix
train %>%
  sample_n(10000) %>%
  ggplot(aes(x = log(winPoints), y = winPlacePerc)) +
  geom_point() +
  labs(x = "Win Points of Player",
       y = "Win Placement Percentage",
       title = "Win Points vs. Placement")

```


